A new simplified manual tour, with application to exploring classification boundaries in high dimensions

### 2022-07-07 paper/paper.Rmd
revised intro and manual tour explanation

<!-- Background -->
From a statistical perspective it is rare to have data that are strictly 3D, and so unlike in most computer graphics applications, the more useful methods for data analysis show projections from an arbitrary dimensional space. These are dynamic data visualizations methods and are collected under the term *tours*. Tours involve views of high-dimensional ($p$) data in low-dimensional ($d$) projections. In his original paper on the grand tour, @As85 provided several algorithms for tour paths that could theoretically show the viewer the data *from all sides*. Prior to Asimov's work, there were numerous preparatory developments including @tukey's PRIM-9. PRIM-9 had user-controlled rotations on coordinate axes, allowing one to manually tour through low-dimensional projections. It is impractical to impossible to steer through all possible projections, unlike Asimov's tours which allows one to quickly see many, many different projections. After Asimov there have been many, many tour developments, as summarized in @lee2021.

<!-- Manual tours -->
One such direction of work develops the ideas from PRIM-9, to provide manual control of a tour. @cook_manual_1997 describes controls for 1D (or 2D) projections, in a 2D (or 3D) manipulation space, allowing the user to select any variable axis, and rotate it into or out of or around the projection through horizontal, vertical, oblique, radial or angular changes in value. @spyrison_spinifex_2020 refines this algorithm and implements them to generate animations.

@@

## The new simpler definition

@@

### Algorithm 1

@@

### Algorithm 2

The problem with Algorithm 1 is that the precise values for $V_m$ cannot be specified because the orthonormalisation will change them. This modification will maintain the components of $V_m$ precisely (Figure \ref{fig:manualsequence} (bottom row)). The algorithm is as follows:

@@

### Algorithm 3

@@

### Potential Algorithm 3

>>>>

<!-- Background -->
From a statistical perspective it is rare to have data that are strictly 3D, and so unlike in most computer graphics applications, the more useful methods for data analysis show projections from an arbitrary dimensional space. These are dynamic data visualizations methods and are collected under the term *tours*. Tours involve views of high-dimensional ($p$) data with low-dimensional ($d$) projections. In his original paper on the grand tour, @As85 provided several algorithms for tour paths that could theoretically show the viewer the data *from all sides*. Prior to Asimov's work, there were numerous preparatory developments including @tukey's PRIM-9. PRIM-9 had user-controlled rotations on coordinate axes, allowing one to manually tour through low-dimensional projections. (A video illustrating the capabilities is available through video library of @ASA22.) Steering through all possible projections is impossible, unlike Asimov's tours which allows one to quickly see many, many different projections. After Asimov there have been many, many tour developments, which are summarized in @lee2021.

<!-- Manual tours -->
One such direction of work develops the ideas from PRIM-9, to provide manual control of a tour. @cook_manual_1997 describe controls for 1D (or 2D) projections, respectively in a 2D (or 3D) manipulation space, allowing the user to select any variable axis, and rotate it into, or out of, or around the projection through horizontal, vertical, oblique, radial or angular changes in value. @spyrison_spinifex_2020 refined this algorithm and implements them to generate animation sequences.

@@

## A new simpler and broadly applicable approach

@@

### Algorithm

@@

## Refinements to enforce exact position

The problem with new simple method (Algorithm 1) is that the precise values for $V_m$ cannot be specified because the orthonormalisation will change them.

### Adjustment method 1

A small modification to algorithm 1 will maintain the components of $V_m$ precisely (Figure \ref{fig:manualsequence} (bottom row)). It is as follows:

@@

### Adjustment method 2

@@

### Adjustment method 3

### 2022-07-07 paper/paper.tex
changed title
revised intro and manual tour explanation

\title{A new simplified manual tour, with examples in mathematica and R}

@@

From a statistical perspective it is rare to have data that are strictly
3D, and so unlike in most computer graphics applications, the more
useful methods for data analysis show projections from an arbitrary
dimensional space. These are dynamic data visualizations methods and are
collected under the term \emph{tours}. Tours involve views of
high-dimensional (\(p\)) data in low-dimensional (\(d\)) projections. In
his original paper on the grand tour, \citet{As85} provided several
algorithms for tour paths that could theoretically show the viewer the
data \emph{from all sides}. Prior to Asimov's work, there were numerous
preparatory developments including \citet{tukey}'s PRIM-9. PRIM-9 had
user-controlled rotations on coordinate axes, allowing one to manually
tour through low-dimensional projections. It is impractical to
impossible to steer through all possible projections, unlike Asimov's
tours which allows one to quickly see many, many different projections.
After Asimov there have been many, many tour developments, as summarized
in \citet{lee2021}.

One such direction of work develops the ideas from PRIM-9, to provide
manual control of a tour. \citet{cook_manual_1997} describes controls
for 1D (or 2D) projections, in a 2D (or 3D) manipulation space, allowing
the user to select any variable axis, and rotate it into or out of or
around the projection through horizontal, vertical, oblique, radial or
angular changes in value. \citet{spyrison_spinifex_2020} refines this
algorithm and implements them to generate animations.

@@

\hypertarget{the-new-simpler-definition}{%
\subsection{The new simpler
definition}\label{the-new-simpler-definition}}

@@

\hypertarget{algorithm-1}{%
\subsubsection{Algorithm 1}\label{algorithm-1}}

@@

\hypertarget{algorithm-2}{%
\subsubsection{Algorithm 2}\label{algorithm-2}}

The problem with Algorithm 1 is that the precise values for \(V_m\)
cannot be specified because the orthonormalisation will change them.
This modification will maintain the components of \(V_m\) precisely
(Figure \ref{fig:manualsequence} (bottom row)). The algorithm is as
follows:

@@

\hypertarget{algorithm-3}{%
\subsubsection{Algorithm 3}\label{algorithm-3}}

@@

\hypertarget{potential-algorithm-3}{%
\subsubsection{Potential Algorithm 3}\label{potential-algorithm-3}}

>>>>

\title{A new simplified manual tour, with application to exploring
classification boundaries in high dimensions}

@@

From a statistical perspective it is rare to have data that are strictly
3D, and so unlike in most computer graphics applications, the more
useful methods for data analysis show projections from an arbitrary
dimensional space. These are dynamic data visualizations methods and are
collected under the term \emph{tours}. Tours involve views of
high-dimensional (\(p\)) data with low-dimensional (\(d\)) projections.
In his original paper on the grand tour, \citet{As85} provided several
algorithms for tour paths that could theoretically show the viewer the
data \emph{from all sides}. Prior to Asimov's work, there were numerous
preparatory developments including \citet{tukey}'s PRIM-9. PRIM-9 had
user-controlled rotations on coordinate axes, allowing one to manually
tour through low-dimensional projections. (A video illustrating the
capabilities is available through video library of \citet{ASA22}.)
Steering through all possible projections is impossible, unlike Asimov's
tours which allows one to quickly see many, many different projections.
After Asimov there have been many, many tour developments, which are
summarized in \citet{lee2021}.

One such direction of work develops the ideas from PRIM-9, to provide
manual control of a tour. \citet{cook_manual_1997} describe controls for
1D (or 2D) projections, respectively in a 2D (or 3D) manipulation space,
allowing the user to select any variable axis, and rotate it into, or
out of, or around the projection through horizontal, vertical, oblique,
radial or angular changes in value. \citet{spyrison_spinifex_2020}
refined this algorithm and implements them to generate animation
sequences.

@@

\hypertarget{a-new-simpler-and-broadly-applicable-approach}{%
\subsection{A new simpler and broadly applicable
approach}\label{a-new-simpler-and-broadly-applicable-approach}}

@@

\hypertarget{algorithm}{%
\subsubsection{Algorithm}\label{algorithm}}

@@

\hypertarget{refinements-to-enforce-exact-position}{%
\subsection{Refinements to enforce exact
position}\label{refinements-to-enforce-exact-position}}

The problem with new simple method (Algorithm 1) is that the precise
values for \(V_m\) cannot be specified because the orthonormalisation
will change them.

\hypertarget{adjustment-method-1}{%
\subsubsection{Adjustment method 1}\label{adjustment-method-1}}

A small modification to algorithm 1 will maintain the components of
\(V_m\) precisely (Figure \ref{fig:manualsequence} (bottom row)). It is
as follows:

@@

\hypertarget{adjustment-method-2}{%
\subsubsection{Adjustment method 2}\label{adjustment-method-2}}

@@

\hypertarget{adjustment-method-3}{%
\subsubsection{Adjustment method 3}\label{adjustment-method-3}}
